Titeln på avhandlingens språk

### 2023-01-30 body/part1/background.tex
edge computing diagram
edge

Through this design, cloud computing affords significant advantages to users.
As services are deployed in a centralized manner accessible over the internet, users can interact with their data and applications from anywhere in the world, from any device.
Thanks the economies of scale and virtualization technologies, the cloud is highly scalable;
services can be scaled simply by spawning more \glspl{VM}.
The specialized nature of cloud providers, the scale of modern datacenters, and the use of virtualization also make the cloud highly reliable.
When hardware fails, recovering is simply a matter of migrating the service container or \gls{VM} to an available compute node~\cite{endo2016high}.

However, the cloud is not suitable for all applications, and presents important drawbacks and challenges for bandwidth-hungry, privacy-requiring, and/or latency-sensitive applications.
Current cloud infrastructures are ill-suited to deal with the tremendous amounts of data generated by \gls{IOT} sensor networks.
The centralized nature of cloud data centers, which are often placed a large distance away from potential users, lead to prohibitively high latencies for highly interactive immersive applications such as mobile \gls{XR}\citationeeded.
Furthermore, this same centralized nature has led to concerns about the security of cloud computing\todo{needs to be expanded a bit}.

Edge computing emerges as an answer to these challenges~\cite{satyanarayanan2009case,satyanarayanan2017emergence}.
This term refers to a novel distributed computing paradigm, complementary to cloud computing, which tackles the challenges of the cloud by moving computation closer to where it is needed.
Instead of offloading computation to a datacenter potentially thousands of kilometers away, in edge computing it is offloaded to a micro-datacenter --- a \emph{cloudlet}~\cite{satyanarayanan2009case} --- at the \emph{edge} of the network, one or two hops away from the user.
This design leads to several advantages.
One, edge computing is uniquely suited for highly latency-sensitive yet resource-intensive applications such as the aforementioned mobile \gls{XR} as well as \gls{CPS}~\cite{sasaki2016vehicle,wang2018bandwidth,wan2020efficient}.

>>>>

Through this design, cloud computing affords significant advantages to users.
As services are deployed in a centralized manner accessible over the internet, users can interact with their data and applications from anywhere in the world, from any device.
Thanks to economies of scale and virtualization technologies, the cloud is highly scalable;
services can be scaled simply by spawning more \glspl{VM}.
The specialized nature of cloud providers, the scale of modern datacenters, and the use of virtualization also make the cloud highly reliable.
When hardware fails, recovering is simply a matter of migrating the service container or \gls{VM} to an available compute node~\cite{endo2016high}.

However, the cloud is not suitable for everything, and presents important drawbacks and challenges for latency-sensitive applications.
In order to achieve the necessary economies of scale, cloud datacenters are designed to serve users distributed across vast geographical areas.
These installations are thus often located ``far'' from potential users;
for instance, at the time of writing, \gls{AWS} processes traffic from all of Scandinavia and the Baltic countries through a single datacenter in Stockholm~\cite{awsregions}.
This leads to prohibitively high latencies for both highly interactive immersive applications such as mobile \gls{XR} and for \glspl{CPS} and \glspl{NCS}~\cite{satyanarayanan2009case,varghese2016challenges,shi2016promise}.
The former category requires \emph{motion-to-photon} latencies (i.e.\ time between input capture and feedback) below \SI{60}{\milli\second} for interactions to be perceived as fluid and responsive by the user~\cite{chen2017empirical}; the latter can require sub-\SI{10}{\milli\second} latencies, for instance in the case of vehicular safety systems.
Such latencies are unfeasible to consistently achieve with cloud computing~\cite{dang2021cloudy}.

Current cloud infrastructures are also ill-suited to deal with the bandwidth requirements of the \gls{IOT} and of mobile computing.
As smart devices, appliances, and sensors become more and more ubiquitous, the network architectures of modern datacenters face significant challenges to deal with the massively increasing volume of traffic~\cite{shi2016edge,wang2019towards}.

\begin{figure}
    \centering
    \includegraphics[height=30em]{Figs/edgecomputing}
    \caption{%
        Conceptual design of edge computing.
        Micro-clouds (known as ``cloudlets''~\cite{satyanarayanan2009case}) are placed a at the edge of the network, a few hops away from end users.
        In other words, these cloudlets are located between users and the internet and the cloud.
    }\label{fig:edgecomputing}
\end{figure}

Edge computing, emerges as a potential answer to these challenges~\cite{satyanarayanan2009case,shi2016promise,shi2016edge,varghese2016challenges,satyanarayanan2017emergence,wang2019towards}.
This term refers to a novel distributed computing paradigm, complementary to cloud computing, which tackles the challenges of the cloud by moving computation closer to where it is needed.
Instead of offloading computation to a datacenter potentially thousands of kilometers away, in edge computing it is offloaded to a micro-datacenter --- a \emph{cloudlet}~\cite{satyanarayanan2009case} --- at the \emph{edge} of the network.
These coudlets feature most of the key characteristics of cloud datacenters, such as multi-tenancy, virtualization of computing resources, virtually unrestricted access to energy, as well as limited scalability, one or two hops away from the user (see \cref{fig:edgecomputing})

The architecture of edge computing offers several advantages.
Thanks to their proximity to users, cloudlets can serve highly latency-sensitive and resource-intensive applications with extremely low latencies while still providing orders of magnitude more computing and energy resources than those present on mobile and wearable devices.
Mobile \gls{XR}, such as \gls{WCA}~\cite{ha2014towards,chen2018application,wang2020scaling,chen2017empirical,chen2018application}, and 5G-enabled \glspl{CPS} and \glspl{NCS}~\cite{sasaki2016vehicle,wang2018bandwidth,wan2020efficient} are uniquely poised to take advantage of this characteristic of edge computing.
